---
title: "project-3"
author: "Justin Feathers"
date: '2022-11-01'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE)
```

# Introduction
Describes the data and variables we want to use. Target is `shares`
```{r}
library(tidyverse)
library(corrplot)
```

# Data
variables of interest: n_tokens_content, num_imgs, num_videos, kw_min_min, kw_max_min,
                       kw_min_max, kw_max_max, is_weekend 
```{r}
newsData <- read_csv(file = "./OnlineNewsPopularity.csv")
techData <- newsData %>% 
            filter(data_channel_is_tech == 1) %>%
            select(n_tokens_content, num_imgs, num_videos, kw_min_min, kw_max_min,
                       kw_min_max, kw_max_max, is_weekend, shares)  
summary(newsData)
summary(techData)
techData
```

# Summarizations
To start, we can look at a correlation plot to see if multicollinearity exists between
any of the variables. Using `corrplot`, we can see that the variables `kw_min_min` and
`kw_max_max` have a strong negative correlation of -0.83, so we will drop `kw_min_min`
from the dataset.

We can analyze a few of the variables by plotting them against `shares`. If we create
a scatter plot of `num_imgs` by `shares`, we can see an outlier when `num_imgs` = 1.
Let's remove that.
```{r}
correlation <- cor(techData)
corrplot(correlation, type = "upper", tl.pos = "lt")
corrplot(correlation, type = "lower", method = "number",
         add = TRUE, diag = FALSE, tl.pos = "n")
techData <- techData %>%
            select(-kw_min_min)
g <- ggplot(techData, aes(x = num_imgs, y = shares))
g + geom_point()
```

We can see the outlier is a single point. We can find the value by using a `summary`
statement. After filtering out outlier, we can see the scatter plot looks much more
reasonable. Based on this plot, it looks like articles with 0 or 1 images tend to
get the most shares with a quadratic decline until hitting the local minimum at 5 images
where it changes to a positive upswing until 11 images. It looks as though `shares`
continues on a negative linear trend after that.
```{r}
summary(techData$shares)
noOutlier <- techData %>%
             filter(shares != 663600)

g <- ggplot(noOutlier, aes(y = shares))
g + geom_point(aes(x = num_imgs))

g + geom_point(aes(x = n_tokens_content))

g + geom_point(aes(x = is_weekend))
```











