---
title: "project-3"
author: "Justin Feathers"
date: '2022-11-01'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE)
```

# Introduction
Describes the data and variables we want to use. Target is `shares`
```{r}
library(tidyverse)
library(corrplot)
library(caret)
```

# Data
```{r}
newsData <- read_csv(file = "./OnlineNewsPopularity.csv")
techData <- newsData %>% 
            filter(data_channel_is_tech == 1) %>%
              select(-url, -timedelta)
```

# Summarizations
A good way of starting is by checking how strongly all variables are correlated to
the response variable of interest. I created a correlation matrix using the `cor`
function and sorted the absolute values of the output to get a convenient tibble
of descending correlation values.

From here, we can look at a correlation plot of the chosen variables to see if 
multicollinearity exists between any of the variables. Using `corrplot`, we can 
see that the variables `n_unique_tokens` and `n_tokens_content` have a strong negative correlation
of -0.73. There is an extremely strong positive correlation of 0.93 between `n_unique_tokens`
and `n_non_stop_unique_tokens` -- we will try dropping `n_unique_tokens` from the dataset
and reassessing. We can see in the new `corrplot` that multicollinearity has been
minimized as desired.

We can analyze a few of the variables by plotting them against `shares`. If we create
a scatter plot of `num_imgs` by `shares`, we can see an outlier when `num_imgs` = 1.
Let's remove that.
```{r}
techCor <- cor(techData$shares, techData) %>%
        as.tibble() %>%
        abs() %>%
        sort(decreasing = TRUE)
techCor

techData <- techData %>% 
            select(num_hrefs, n_tokens_content, n_unique_tokens, kw_avg_avg,
                   n_non_stop_unique_tokens, num_videos, min_negative_polarity,
                   rate_positive_words, min_positive_polarity, shares)

correlation <- cor(techData)
corrplot(correlation, type = "upper", tl.pos = "lt")
corrplot(correlation, type = "lower", method = "number",
         add = TRUE, diag = FALSE, tl.pos = "n")

techData <- techData %>%
            select(-n_unique_tokens)

corrplot(correlation, type = "upper", tl.pos = "lt")
corrplot(correlation, type = "lower", method = "number",
         add = TRUE, diag = FALSE, tl.pos = "n")
```

We can see the outlier is a single point. We can find the value by using a `summary`
statement. After filtering out the outlier, we can see the scatter plot looks much more
reasonable. Based on this plot, it looks like articles with 0 or 1 images tend to
get the most shares with a quadratic decline until hitting the local minimum at 5 images
where it changes to a positive upswing until 11 images. It looks as though `shares`
continues on a negative linear trend after that. Inspecting the plot for `shares` 
vs. `n_tokens_content` seems to suggest shares tend to decrease after articles go 
beyond 250-500 words. Next, the plot of `shares` vs. `rate_positive_words` suggests 
that articles are far more likely to be shared as the rate of positive words increases.
Finally, we can see quartiles and means for the variables using the `summary` function
and standard deviations with the `sd` function.
```{r}
summary(techData$shares)
noOutlier <- techData %>%
             filter(shares != 663600)

g <- ggplot(noOutlier, aes(y = shares))
g + geom_point(aes(x = num_hrefs))
g + geom_point(aes(x = n_tokens_content))
g + geom_point(aes(x = rate_positive_words))

summary(techData)
techData %>%
  sapply(sd)
```

# Modeling
```{r}
set.seed(5000)
techIndex <- createDataPartition(techData$shares, p = 0.70, list = FALSE)
techTrain <- techData[techIndex, ]
techTest <- techData[-techIndex, ]
```

## Multiple Linear Regression
Fitting a multiple regression model on all variables in the `techTrain` dataset,
we can see from the `summary` function that this model is not a very good fit with
an adjusted R^2 value of 0.01 -- this means only 1% of the variance in the data is
explained by the model. We need to explore better options.

```{r}
mlrFit <- train(shares ~ ., data = techTrain,
                preProcess = c("center", "scale"),
                method = "lm")
mlrPredict <- predict(mlrFit, newdata = techTest)
postResample(mlrPredict, techTest$shares)
```

## Random Forest
```{r}
techForest <- train(shares ~ ., data = techTrain,
                method = "rf",
                preProcess = c("center", "scale"),
                tuneGrid = data.frame(mtry = ncol(techTrain)/3),
                trControl = trainControl(method = "cv", number = 5))
forestPredict <- predict(techForest, newdata = techTest)
postResample(forestPredict, techTest$shares)
```












